Binning Strategies and Related Loss for Large Data

### 2015-04-12 Paper.Rnw
Add brief discussion of unit interpretability of net spatial loss if scales and/or units differ between the two  binned dimensions.

% The net spatial loss is a euclidean distance, but requires a reference scale on which the magnitude may be interpretable. We  use the spatial loss from a ``worst-case'' binning scenario as a reference value. The largest displacement of points  occurs if we place all points into a single bin.  While the total spatial loss for single bin could be made arbitrarily large if the bin center is placed far away from the data values, we will consider a reasonable specification of a bin with boundaries that span only the range of the data in each dimension centered at  $( (x_{(1)}-x_{(n)})/2$ , $(y_{(1)}-y_{(n)})/2 )$.  The total spatial loss for this ``worst-case'' bin will be referred to as the \textit{potential spatial loss}, $L^S_p$, which is defined as 
% \begin{eqnarray}\label{potspatloss}
% L^S_p = \sum_{i=1}^{n} \sqrt{ \left(x_i-(x_{(1)}-x_{(n)})/2\right)^2 + \left(y_i-(y_{(1)}-y_{(n)})/2\right)^2 }
% \end{eqnarray} 
% %
% We use net spatial loss as a proportion of potential spatial loss, $L^S_{net}/L^S_p$,  as the primary metric for assessing spatial loss incurred from binning.

>>>>

\ktm{The net spatial loss is a euclidean distance, but the units affiliated with this distance are based on the units on which each variable is recorded.  If the two variables in the binned scatterplot share the same units this leads to direct interpretability of the spatial loss.  However, if the two variables do not share the same units or magnitude of values it is advisable to standardize the variables prior to binning, thus the spatial loss is more universally interpretable as a distance in units of standard devations. }

%but requires a reference scale on which the magnitude may be interpretable. We  use the spatial loss from a ``worst-case'' binning scenario as a reference value. The largest displacement of points  occurs if we place all points into a single bin.  While the total spatial loss for single bin could be made arbitrarily large if the bin center is placed far away from the data values, we will consider a reasonable specification of a bin with boundaries that span only the range of the data in each dimension centered at  $( (x_{(1)}-x_{(n)})/2$ , $(y_{(1)}-y_{(n)})/2 )$.  The total spatial loss for this ``worst-case'' bin will be referred to as the \textit{potential spatial loss}, $L^S_p$, which is defined as 
% \begin{eqnarray}\label{potspatloss}
% L^S_p = \sum_{i=1}^{n} \sqrt{ \left(x_i-(x_{(1)}-x_{(n)})/2\right)^2 + \left(y_i-(y_{(1)}-y_{(n)})/2\right)^2 }
% \end{eqnarray} 
% %
% We use net spatial loss as a proportion of potential spatial loss, $L^S_{net}/L^S_p$,  as the primary metric for assessing spatial loss incurred from binning.
